### 2025-01-09 skripsi/apendix.tex
finish ch2 and start ch5

% !TEX root = ./skripsi.tex
\chapter{KAJIAN PUSTAKA}

\noindent Bab ini mengulas secara rinci konsep-konsep dasar yang berkaitan dengn pekerjaan penelitian TA dan deskripsi studi pustaka yang dilakukan. Judul bab tidak harus seperti yang dituliskan, melainkan dapat lebih fleksibel yang mencerminkan isi paparan pada bab ini. Demikian halnya dengan judul sub bab.

\section{Fourier Transform and Series}\label{sec:fourier_discussion}
\noindent

Complex numbers are numbers consisting of real and imaginary components. Complex numbers are written as in \lccref{eq:complex_number} with \(a \) and \(b \) being the real and imaginary components respectively. The imaginary component is multiplied by the imaginary unit \(i=\sqrt{-1}\).
\begin{align}
  c = a + bi \label{eq:complex_number} \\
\end{align}
\autocite{DiscreteFourierTransform}
\begin{align}
  c_k=\sum_{n=0}^{N-1} f_n e^{-i2\pi kn/N} \label{eq:discrete_fourier_transform} \\
  f_n=\sum_{k=0}^{N-1} c_k e^{i2\pi kn/N} \label{eq:inverse_discrete_fourier_transform}
\end{align}

\section{Least Squares Support Vector Machine}
\noindent An arguably fundamental model widely used whether in pedagogical settings or otherwise is the support vector machine. It dates back to works by Vapnik \& Lerner in 1963 (Recognition of Patterns with help of Generalized Portraits) and V. N. Vapnik \& A. Ya. Chervonenkis in 1964 (A note on one class of perceptrons/On a perceptron class). As the development on SVM continued, what originally was a model for classification of separable data generalized to regression tasks as well Vapnik (2000 The Nature of Statistical Learning Theory).
\subsection{Disadvantages}
However, the main disadvantage of LSSVMs are the fact that they do not have the sparse property of SVMs which can leave performance on the table. A simple mitigation can be done by filtering training samples with small absolute values of lagrangian multipliers \autocite{haifengwangComparisonSVMLSSVM2005}.

To derive the least squares support vector regression (J.A. Suykens LSSVM Book) model we start with the linear expression:

\begin{equation}
  y=W^{\intercal}\vb{x}+b
\end{equation}

\begin{equation}
  \min_{W,e} J(W,e) = \frac{1}{2}W^{\intercal}W + C\frac{1}{2}\sum_{k=1}^{n}e_{k}
\end{equation}

Such that
\begin{equation}
  y_{k}=W^{\intercal}\vb{x}_{k}+b+e_{k} \qquad k=1,\dots,n
\end{equation}

\begin{equation}
  L(W,b,e;\vb{\alpha}) = \frac{1}{2}W^{\intercal}W + C\frac{1}{2}\sum_{k=1}^{n}e_{k} + \sum_{k=1}^{n}\alpha_{k} \left(W^{\intercal}\vb{x}_{k}+b+e_{k} - y_{k} \right)
\end{equation}

Derive the KKT system
\begin{equation}
  \begin{split}
    \pdv{L}{W} = 0 \rightarrow          & W = \sum_{k=1}^{n}\alpha_{k}x_{k}                                \\
    \pdv{L}{b} = 0 \rightarrow          & \sum_{k=1}^{n}\alpha_{k} = 0                                     \\
    \pdv{L}{e_{k}} = 0 \rightarrow      & \alpha_{k} = C e_{k}                        & \qquad k=1,\dots,n \\
    \pdv{L}{\alpha_{k}} = 0 \rightarrow & W^{\intercal}\vb{x}_{k}+b+e_{k} - y_{k} = 0 & \qquad k=1,\dots,n \\
  \end{split}
\end{equation}

After eliminating $W$ and $\vb*{e}$, with $\vb{1}_{n} = \langle 1,\dots,1\rangle$, $\vb{y} = \lbrack y_1, \dots, y_n \rbrack $, and ${\alpha} = \lbrack \alpha_1, \dots, \alpha_n \rbrack $ the solution is as follows in block matrix notation
\begin{equation} \label{eq:lssvr_solution}
  \begin{bmatrix}
    0          & \vb{1}_{n}^{\intercal} \\
    \vb{1}_{n} & \Omega + \frac{I}{C}
  \end{bmatrix}
  \begin{bmatrix}
    b           \\
    \vb{\alpha} \\
  \end{bmatrix}
  =
  \begin{bmatrix}
    0      \\
    \vb{y} \\
  \end{bmatrix}
\end{equation}

The solution in \cref{eq:lssvr_solution}

\begin{algorithm}[H]
  \caption{LSSVR Training}\label{alg:lssvr_training}
  \begin{algorithmic}[1]
    \Procedure{SolveLSSVR}{$\vb{X}, \vb{y}, \gamma $}
    \State{\(\Omega \gets [[]]\)}
    \Comment{Construct matrix of inner products in high dimensional space}

    \For{\(k=0 \to n\)}
    \For{\(l=0 \to n\)}
    \State{\(\Omega_{k,l}\gets K(\vb{X}_k,\vb{X}_l)\)}
    \EndFor{}
    \EndFor{}
    \State{\(\vb{H} \gets \Omega + \frac{\vb{I}}{\gamma}\)}
    % \State $\vb{1} \gets [1,1,...,1]$
    \State{\(\vb{A} \gets [[]]\)}
    \Comment{Construct left-hand side matrix}
    \State{\(\vb{A}_{0,0} \gets 0\)}
    \For{\(k=0 \to n\)}
    \State{\(\vb{A}_{k+1,0} \gets 1\)}
    \State{\(\vb{A}_{0,k+1} \gets 1\)}
    \EndFor{}
    \For{\(k=0 \to n\)}
    \For{\(l=0 \to n\)}
    \State{\(\vb{A}_{k+1,l+1} \gets \vb{H}_{k,l}\)}
    \EndFor{}
    \EndFor{}
    \State{\(\vb{B}\gets []\)}
    \Comment{Construct left-hand side of the equation}
    \State{\(\vb{B}_0 \gets 0\)}
    \For{\(k=0 \to n\)}
    \State{\(\vb{B}_{k+1} \gets \vb{y}_{k}\)}
    \EndFor{}
    \State{\(\vb{A}^{\dagger} \gets pseudoInverse(\vb{A})\)}
    \Comment{Compute solution using pseudo inverse}
    \State{\(\vb{S} \gets \vb{A}^{\dagger}\vb{B}\)}
    \State{\(b \gets \vb{S}_{0}\)}
    \For{\(k=0 \to n\)}
    \State{\(\vb{\alpha}_{k} \gets \vb{S}_{k+1}\)}
    \EndFor{}
    \State{\textbf{return}\(\vb{\alpha}, b\)}
    \EndProcedure{}
  \end{algorithmic}
\end{algorithm}

The basic pseudocode from the LSSVM Equation for function regression is defined in \lccref{alg:lssvr_training} for a training set of length $n$, features $\vb{X}$, and labels $\vb{y}$. Training the LSSVM means computing the values of langrange multipliers $\vb{\alpha}$ and bias $b$. $K$ is the kernel function used to compute the inner products in high dimensional space, here we assume the RBF kernel. $\vb{A}$ is a matrix of size $n+1$ by $n+1$. $\vb{H}$ is a matrix of size $n$ by $n$. $\vb{I}$ is the identity. $\vb{B}$ is a vector of size $n+1$. $\vb{S}$ is a vector of size $n+1$.

After training the model can be used for prediction of unseen features. The pseudocode for prediction is shown in \lccref{alg:lssvr_prediction} for prediction features $\vb{U}$ with \(p \) samples. The trained model uses the training features themselves $\vb{X}$ with \(n \) samples, the learned multipliers of training points $\vb{\alpha}$, and the bias $b$.

\begin{algorithm}[H]
  \caption{LSSVR Prediction}\label{alg:lssvr_prediction}
  \begin{algorithmic}[1]
    \Require $\vb{U}, \vb{\alpha}, \vb{X}, b$
    \Ensure $\vb{v}$
    \State $\Omega \gets [[]]$ \Comment{Construct matrix of inner products in high dimensional space}

    \For{$k=0 \to p$}
    \For{$l=0 \to n$}
    \State$\Omega_{k,l}\gets K(\vb{U}_{k},\vb{X}_l)$
    \EndFor
    \EndFor
    \State $\vb{v} \gets \Omega\vb{\alpha} + \vb{1}_{m}b$
  \end{algorithmic}
\end{algorithm}

In this example we will be using the function \(2x^2+4\). The values of this function can be seen in \lccref{table:1}.

\begin{table}[H]
  \centering
  \begin{tabular}{@{}lll@{}}
    \toprule
    No & x       & y       \\ \midrule
    1  & 0.0     & 4.0     \\
    2  & 0.33... & 4.22... \\
    3  & 0.66... & 4.88... \\
    4  & 1.0     & 6.0     \\
    \bottomrule
  \end{tabular}
  \caption{Example data of function $2x^2+4$} \label{table:1}
\end{table}
For $\Omega_{i,j}=K(x_i,x_j)=exp \left(-\frac{\norm*{x_i-x_j}^2}{2\sigma^2}\right)$

For example, with $\sigma=1$, $x_i=0.0$, \& $x_j = 0.33\dots$
\begin{equation}
  \begin{split}
    K(0.0,0.33) & = exp\left(-\frac{\norm*{0.0-0.33}^2}{2(1)^2}\right) \\
                & = exp\left(-\frac{0.33^2}{2}\right)                  \\
                & = 0.9460
  \end{split}
\end{equation}

@@

\begin{equation}
  \vb{I}\frac{1}{\gamma}\to\vb{I}\frac{1}{5}\to
  \begin{bmatrix}
    0.2000 & 0.0000 & 0.0000 & 0.0000 \\
    0.0000 & 0.2000 & 0.0000 & 0.0000 \\
    0.0000 & 0.0000 & 0.2000 & 0.0000 \\
    0.0000 & 0.0000 & 0.0000 & 0.2000
  \end{bmatrix}
\end{equation}

@@

\begin{equation}
  \Omega\alpha+b\vb{1}_{m}\to v\to
  \begin{bmatrix}
    4.4516 \\
    4.3251 \\
    4.7413 \\
  \end{bmatrix}
\end{equation}
Where $\vb{1}_{m}$ is a vector of 1s with the length of $U$.

% \section{Sub Bab B}
% \noindent Suatu penelitian tidak dapat lepas dari capaian pengetahuan dan pemahaman yang sudah dipublikasikan. Deskripsi tentang capaian ini menjadi penting karena selain menunjukkan tingkat pemahaman mahasiswa, juga mengetahui tempat pekerjaan penelitian TA dalam konstelasi capaian tersebut. Studi pustaka dan paparan hasilnya dapat memperkaya wawasan tentang topik yang diangkat pada penelitian TA.

% \section{Membuat Persamaan}
% \noindent Secara prinsip, suatu persamaan menyatu dalam kalimat. Letak persamaan dapat berada di awal, tengah, atau akhir kalimat. Dengan demikian, pada akhir persamaan harus diberikan tanda baca, misalnya koma, titik koma, atau titik, yang menekankan kehadiran persamaan dalam kalimat. Tidak semua persamaan harus diberi nomor. Persamaan yang dirujuk pada naskah TA saja yang harus diberi nomor. Kode awal penomoran ini adalah nomor urut bab, termasuk untuk persamaan pada Lampiran, dengan urutan alfabet kapital.

% Setiap notasi harus unik atau tunggal, sehingga arti setiap notasi adalah unik atau tunggal juga. Arti satu notasi harus dituliskan segera ketika notasi tersebut muncul, dan tidak diulang lagi setelahnya.

% \subsection{Contoh Persamaan Sederhana}
% Persamaan (\ref{II.1}) mendeskripsikan dinamika fungsi gelombang $\psi(\vec{r},t)$ di bawah pengaruh potensial $V(\vec{r})$ dan dituliskan sebagai berikut:
% \begin{equation}\label{II.1}
%   i\hslash\frac{\partial \psi}{\partial t} = -\frac{\hslash^2}{2 m}\Vec{\nabla}^2 \psi + V(\vec{r})\psi,
% \end{equation}
% dengan $m$ adalah massa partikel dan $\hslash$ merupakan konstanta Planck tereduksi.

% Di akhir persamaan (\ref{II.1}) diberi koma karena berada ditengah kalimat. Untuk merujuk ke persamaan yang telah ditulis, gunakan perintah \verb|\ref{}|.

% Untuk menuliskan beberapa set persamaan yang masih terhubung, gunakan \verb|\subequations{}|. Misal kita punya persamaan diferensial terkopel, kita bisa tuliskan
% \begin{subequations}
%   \begin{align}
%     \frac{dy}{dt} & = -x , \\
%     \frac{dx}{dt} & = -y .
%   \end{align}
% \end{subequations}
% Kalau perlu matriks, kita bisa tulis seperti berikut.
% \begin{equation}
%   \sigma_x =
%   \begin{pmatrix}
%     0 & 1 \\
%     1 & 0
%   \end{pmatrix},\quad
%   \sigma_y =
%   \begin{pmatrix}
%     0 & -i \\
%     i & 0
%   \end{pmatrix}, \quad
%   \sigma_z =
%   \begin{pmatrix}
%     1 & 0  \\
%     0 & -1
%   \end{pmatrix}
% \end{equation}

% \section{Referensi dan Citation}
% \noindent Sitasi dapat dimasukkan ke dalam Tugas Akhir seperti ini \cite{Fujita1996}. Untuk sitasi dengan beberapa sumber, dapat dituliskan juga \cite{hohen1964,Kim2006}. Atau untuk tiga sumber berarti \cite{kongkanand2006,kresse1999,Leibb1993}.

>>>>

% !TEX root = ./skripsi.tex
\appendix

\chapter{EXAMPLE COMPUTATION OF LSSVR}
% Example computation
In this example we will be using the function \(2x^2+4\). The values of this function can be seen in \lccref{table:1}.

\begin{table}[H]
  \centering
  \begin{tabular}{@{}lll@{}}
    \toprule
    No & x          & y          \\ \midrule
    1  & 0.0        & 4.0        \\
    2  & 0.33\ldots & 4.22\ldots \\
    3  & 0.66\ldots & 4.88\ldots \\
    4  & 1.0        & 6.0        \\
    \bottomrule
  \end{tabular}
  \caption{Example data of function \(2x^2+4\)}\label{table:lssvr_example_function_values}
\end{table}
For \(\Omega_{i,j}=K(x_i,x_j)=exp \left(-\frac{\norm*{x_i-x_j}^2}{2\sigma^2}\right)\)

For example, with \(\sigma=1\), \(x_i=0.0\), \& \(x_j = 0.33\dots\)
\begin{equation}
  \begin{split}
    K(0.0,0.33) & = \exp\left(-\frac{\norm*{0.0-0.33}^2}{2(1)^2}\right) \\
                & = \exp\left(-\frac{0.33^2}{2}\right)                  \\
                & = 0.9460
  \end{split}
\end{equation}

@@

\begin{equation}
  \vb{I}\frac{1}{C}\to\vb{I}\frac{1}{5}\to
  \begin{bmatrix}
    0.2000 & 0.0000 & 0.0000 & 0.0000 \\
    0.0000 & 0.2000 & 0.0000 & 0.0000 \\
    0.0000 & 0.0000 & 0.2000 & 0.0000 \\
    0.0000 & 0.0000 & 0.0000 & 0.2000
  \end{bmatrix}
\end{equation}

@@

\begin{equation}
  \Omega\alpha+b\vb{1}_{m}\to v\to
  \begin{bmatrix}
    4.4516 \\
    4.3251 \\
    4.7413 \\
  \end{bmatrix}
\end{equation}
Where \(\vb{1}_{m}\) is a vector of 1s with the length of \(U\).

% \section{PROGRAM SATU}
% \section{PROGRAM DUA}

% \chapter{GAMBAR-GAMBAR}

### 2025-01-09 skripsi/chapter1.tex
finish ch2 and start ch5

Forecasting the weather or evolution of any other system is formulated as an initial value problem where past observations are used to predict an unknown future state of the system. A different formulation can be how heat spreads over time in different types of frying pans from known information of the pan's materials, their heat conduction properties, and that the stove gives off constant heat. In general the scenario is predicting the effect (future atmospheric temperature distribution, etc.) of some given causes (current and past weather, etc.). This scenario is termed the forward problem. The inverse problem on the other hand solves for an unknown parameter using other known parameters and observations of the partial solution such as temperature field at the object's surface. Traditional methods used to solve these problems utilized knowledge of the exact equations to numerically solve PDEs. As an example, the Finite Difference Method (FDM) approximates the solution by substituting the partial derivatives with finite differences and manipulating the equation such that the solution can be computed. The inverse problem on the other hand is determination of initial conditions or other parameters from observed state of the system. For example determining the heat source distribution given the temperature distribution at some point in time. One traditional approach to this would have meant solving the forward problem from an initial guess of the parameter and then evaluating some cost function such as how close the predicted solution is to the observed one. As expected, this would require many evaluations of the forward solver. This means that the cost of the solver would be greatly multiplied, emphasizing the importance of an efficient solver.
% TODO: reword this because the explanation of the problem setup for forward and inverse problems is not layed out optimally.
% TODO: perhaps the disadvantage of inverse problem in the paragraph above should be removed and only explanation of the inverse problem setup itself left by itself
% \begin{equation} \label{eq:forward_finite_difference_1}
%     \pdv{f\left(x\right)}{x} \approx \frac{f\left(x+h\right)-f\left(x\right)}{h}
% \end{equation}

@@

\subsubsection*{Operator Learning for Weather Forecasting}
In terms of weather forecasting, operator learning has been applied in terms of initial value problems. This problem formulation is reminiscent of time series prediction problems widely found in machine learning research. Researchers \textcite{kurthFourCastNetAcceleratingGlobal2023} developed FourCastNet which utilized Adaptive Fourier Neural Operator (AFNO), a transformer based model containing the previously mentioned FNO computational blocks by \textcite{li2021fourier}. This model was then able to be trained in a massively parallel manner. In a comparison with a traditional model called the Integrated Forecasting System from the European Center for Medium Range Weather Forecasts (ECMWF), FourCastNet is faster and much more efficient in terms of inference time, resulting in about 80,000 times speed up for a 100-member ensemble forecast. This is while performing much better than a previous deep learning approach. In another study, \textcite{bonevSphericalFourierNeural2023} proposed a variation on neural operators called Spherical Fourier Neural Operator (SFNO) which exploited the spherical nature of global forecasting by using Spherical Harmonic Transform (SHT) in place of Fourier Transform. This model when compared to AFNO and FNO, produced no visible artifacts in autoregressive rollouts for long range forecasting. In terms of forecasting, the model shows outcomes that matches the IFS which is a big leap forward in parity for traditional and machine learning based methods.

>>>>

Forecasting the weather or evolution of any other system is formulated as an initial value problem where past observations are used to predict an unknown future state of the system. A different formulation can be how heat spreads over time in different types of frying pans from known information of the pan's materials, their heat conduction properties, and that the stove gives off constant heat. In general the scenario is predicting the effect (future atmospheric temperature distribution, etc.) of some given causes (current and past weather, etc.). This scenario is termed the forward problem. The inverse problem on the other hand solves for an unknown parameter using other known parameters and observations of the partial solution such as temperature field at the object's surface. Traditional methods used to solve these problems utilized knowledge of the exact equations to numerically solve PDEs. As an example, the Finite Difference Method (FDM) approximates the solution by substituting the partial derivatives with finite differences and manipulating the equation such that the solution can be computed. The inverse problem on the other hand is determination of initial conditions or other parameters from observed state of the system. For example determining the heat source distribution given the temperature distribution at some point in time. One traditional approach to this would have meant solving the forward problem from an initial guess of the parameter and then evaluating some cost function such as how close the predicted solution is to the observed one. As expected, this would require many evaluations of the forward solver. This means that the cost of the solver would be greatly multiplied, emphasizing the importance of an efficient solver.
% TODO: reword this because the explanation of the problem setup for forward and inverse problems is not layed out optimally.
% \begin{equation} \label{eq:forward_finite_difference_1}
%     \pdv{f\left(x\right)}{x} \approx \frac{f\left(x+h\right)-f\left(x\right)}{h}
% \end{equation}

@@

\subsubsection*{Operator Learning for Weather Forecasting}
In terms of weather forecasting, operator learning has been applied in terms of initial value problems. This problem formulation is reminiscent of time series prediction problems widely found in machine learning research. Researchers \textcite{kurthFourCastNetAcceleratingGlobal2023} developed FourCastNet which utilized Adaptive Fourier Neural Operator (AFNO), a transformer based model containing the previously mentioned FNO computational blocks by \textcite{li2021fourier}. This model was then able to be trained in a massively parallel manner. In a comparison with a traditional model called the Integrated Forecasting System from the European Center for Medium Range Weather Forecasts (ECMWF), FourCastNet is faster and much more efficient in terms of inference time, resulting in about \num{80000} times speed up for a \num{100}-member ensemble forecast. This is while performing much better than a previous deep learning approach. In another study, \textcite{bonevSphericalFourierNeural2023} proposed a variation on neural operators called Spherical Fourier Neural Operator (SFNO) which exploited the spherical nature of global forecasting by using Spherical Harmonic Transform (SHT) in place of Fourier Transform. This model when compared to AFNO and FNO, produced no visible artifacts in autoregressive rollouts for long range forecasting. In terms of forecasting, the model shows outcomes that matches the IFS which is a big leap forward in parity for traditional and machine learning based methods.
